---
title: 'put the title of your project here'
author: "put your name here"
date: "Due Oct. 30, 2017"
bibliography: "decarbonization.bib" # this line tells RMarkdown which file to use for the bibliography
output: 
  html_document: default
  pdf_document:
    include:
      in_header: "ees3310.sty" # This line just tells RMarkdown to use a style file to format the PDF file with Times Roman font.
---
```{r setup, include=FALSE}
knitr::knit_hooks$set(inline = function(x) { knitr:::format_sci(x, 'md')})
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)

# This section loads necessary R libraries and sources scripts that define 
# useful functions format_md.
# 
data_dir = "data"
script_dir = "scripts"

if (!dir.exists('data')) dir.create('data')

library(pacman)

# Some lab computers fail to load tidyverse
# because bindr is not installed.
# This will prevent such errors.
if (!p_isinstalled(bindr)) p_install(bindr)


# Load all of the following packages.
# Install any missing packages.
p_load(scales, tidyverse, stringr, janitor, knitr)

# This sets the default style for ggplot
theme_set(theme_bw(base_size = 15))

# Load basic utility scripts
source('scripts/utils.R', chdir = T)

```

# Load World Data and Plot Kaya variables

```{r load_world}
world_data = read_csv('data/World.csv')
```

```{r plot_world_P}
ggplot(world_data, aes(x = year, y= P)) +
  geom_line() + geom_point()
```

```{r plot_world_g}
ggplot(world_data, aes(x = year, y= g)) +
  geom_line() + geom_point()
```

```{r plot_world_e}
ggplot(world_data, aes(x = year, y= e)) +
  geom_line() + geom_point()
```

```{r plot_world_f}
ggplot(world_data, aes(x = year, y= f)) +
  geom_line() + geom_point()
```

```{r plot_world_log_P}
ggplot(world_data, aes(x = year, y= P)) +
  geom_line() + geom_point() +
  scale_y_log10()
```

```{r plot_world_log_g}
ggplot(world_data, aes(x = year, y= g)) +
  geom_line() + geom_point() +
  scale_y_log10()
```

```{r plot_world_log_e}
ggplot(world_data, aes(x = year, y= e)) +
  geom_line() + geom_point() +
  scale_y_log10()
```

```{r plot_world_log_f}
ggplot(world_data, aes(x = year, y= f)) +
  geom_line() + geom_point() + 
  scale_y_log10()
```


```{r calculate_slopes}
# average annual growth rate of per-capita GDP, in percent per year
P_fit = lm(log(P)~ year, data = world_data)
rate_P = summary(P_fit)$coefficients['year', 'Estimate']

g_fit = lm(log(g)~ year, data = world_data)
rate_g = summary(g_fit)$coefficients['year', 'Estimate']

e_fit = lm(log(e)~ year, data = world_data)
rate_e = summary(e_fit)$coefficients['year', 'Estimate']

f_fit = lm(log(f)~ year, data = world_data)
rate_f = summary(f_fit)$coefficients['year', 'Estimate']

rates = tibble(variable = c("P", "g", "e", "f"), 
               rate = c(rate_P, rate_g, rate_e, rate_f))

kable(rates)

```


Now let's calcualte the implied rate of reduction of F: the total CO~2~ emissions?

```{r emissions_reduction_rate}
F_2005 = filter(world_data, year == 2005)$F
F_2050 = F_2005 * (1.0 - 0.36)
F_2016 = filter(world_data, year == 2016)$F

implied_rate_F = log(F_2050 / F_2016) / (2050 - 2016)
```

The policy calls for reducing F so that F in 2050 is 36% less than it was in 2005.
This implies that between 2016 and 2050, we have to reduce F at an average rate of
`r format_md(implied_rate_F * 100, 2)`% per year.


Now let's calcualte the implied rate of decarbonization of the economy.

```{r calculate_implied_decarbonization}
implied_rate_ef = implied_rate_F - (rate_P + rate_g)
```
The implied rate of decarbonization of the economy is
`r format_md(100 * implied_rate_ef, 2)`% per year.
The actual historical  rate of decarbonization is
`r format_md(100 * (rate_e + rate_f), 2)`% per year.
